---
title: "p8105_midterm_mk4022"
author: "Mary Knoop"
date: "10/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load librarires for analysis 
```{r}
devtools::install_github("thomasp85/patchwork")

library(tidyverse)
library(knitr)
library(readxl)
library(dplyr)
library(patchwork)
```

Problem 1
Part 1:
Import and clean the data. Format the data to use appropriate variable names; fill in missing values with data where appropriate (as indicated in the header information); create character and ordered factors for categorical variables.
```{r}
posture_data = 
  read_excel("/Users/maryknoop/Desktop/Data_Sci/p8105_midterm_mk4022/data/p8105_mtp_data.xlsx", skip = 8) %>%
  janitor::clean_names() %>%
  mutate(
    eop_size_mm = replace_na(eop_size_mm, 0),
    eop_size = factor(eop_size, c("0", "1", "2", "3", "4", "5"), c("0-5", "5-10", "10-15", "15-20", "20-25", "25+"), ordered = TRUE), 
    sex = factor(sex, c(0, 1), c("female", "male")),
    fhp_category = factor(fhp_category,c(0, 1, 2, 3, 4, 5, 6, 7),c("0-10", "10-20", "20-30", "30-40", "40-50", "50-60", "60-70", "70-80"), ordered = TRUE),
    age_group = factor(age_group, c("2", "3", "4", "5", "6", "7", "8"), c("18-30", "31-40", "41-50", "51-60", "60+", "60+", "60+"), ordered = TRUE)) %>%
    filter(age_group != "1")
posture_data
```
Part 2:
Briefly describe the data cleaning process and the resulting dataset, identifying key variables based on your understanding of the original scientific report. How many participants are included? What is the age and gender distribution (a human-readable table may help here)?

The dataset posture_data is comprised of 9 columns. Key variables include  We have 1,219 participants included in the posture_data dataset.  The age and gender distribution is
```{r}
agesex_table = posture_data %>%
filter(age_group != "1") %>%
group_by(sex, age_group) %>%
summarize (n= n()) %>%
pivot_wider(
names_from = sex, 
values_from = n
) %>%
knitr::kable ()
agesex_table

agesex_plot=
posture_data %>% 
ggplot(aes(x = age, fill = sex)) +
geom_histogram()
agesex_plot
```

Problem 2
Figure 3 shows only the mean and standard deviation for FHP, but does not show the distribution of the underlying data. Figure 4 shows the number of participants in each age and sex group who have an enlarged EOP (based on categorical EOP Size – groups 0 and 1 vs groups 2, 3, 4, and 5). However, the number of participants in each age and sex group was controlled by the researchers, so the number with enlarged EOP in each group is not as informative as the rate of enlarged EOP in each group. Create a two-panel figure that contains improved versions of both of these.
```{r}
figure3 = 
ggplot(posture_data, aes(x = age, y = fhp_size_mm, na.rm = TRUE)) +
geom_point(aes(color = age_group)) + 
geom_smooth(size = .7, se = TRUE) + 
facet_grid(~sex)
figure3

figure4 = 
posture_data %>% 
mutate(enlarged_eop = as.numeric(ifelse(eop_size_mm >=10, "1", "0"))) %>% 
group_by(age_group, sex) %>% 
summarize(rate = mean(enlarged_eop)) %>% 
ggplot(aes(x = age_group, y = rate, group =sex, color = sex)) +
geom_point () +
geom_line () +
labs(
title = "Rate of enlarged EOP by Age and Sex",
x = "Age_Group",
y = "EOP Rate")
figure4

figure3+figure4
```

Part 2:
Create a 2 x 5 collection of panels, which show the association between FHP size and EOP size in each age and sex group. Comment on your plots with respect to the scientific question of interest.
```{r}
posture_data %>%
  ggplot(aes(x = eop_size_mm, y = fhp_size_mm, na.rm = TRUE, color = age_group)) +
         geom_point () + facet_grid (sex ~ age_group) + labs (title = "2x5 Panel" , 
    x = "eop_size_mm" , 
    y = "fhp_size_mm")

```

Problem 3
Part 1:
* Are the authors’ stated sample sizes in each age group consistent with the data you have available?
* Are the reported mean and standard deviations for FHP size consistent with the data you have available?
* The authors find “the prevalence of EEOP to be 33% of the study population”. What is the definition of EEOP, and what variables can you use to evaluate this claim? Is the finding consistent with the data available to you?
* FHP is noted to be more common in older subjects, with “FHP >40 mm observed frequently (34.5%) in the over 60s cases”. Are the broad trends and specific values consistent with your data?

Sample sizes from paper: 18–30 n=300, 31–40 n=200, 41–50 n=200, 51–60 n=200 and >60 n=300
```{r}
posture_data %>%
  group_by(age_group) %>%
  summarize(n_obs = n())
```
```{r}
mean(pull(filter(posture_data,sex == "female"), fhp_size_mm), na.rm = TRUE)
sd(pull(filter(posture_data,sex == "female"), fhp_size_mm), na.rm = TRUE)
mean(pull(filter(posture_data,sex == "male"), fhp_size_mm), na.rm = TRUE)
sd(pull(filter(posture_data,sex == "male"), fhp_size_mm), na.rm = TRUE)
```



