---
title: "p8105_midterm_mk4022"
author: "Mary Knoop"
date: "10/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load librarires for analysis 
```{r}
devtools::install_github("thomasp85/patchwork")

library(tidyverse)
library(knitr)
library(readxl)
library(dplyr)
library(patchwork)
```

Problem 1
Part 1:
```{r}
posture_data = 
  read_excel("/Users/maryknoop/Desktop/Data_Sci/p8105_midterm_mk4022/data/p8105_mtp_data.xlsx", skip = 8) %>%
  janitor::clean_names() %>%
  mutate(
    eop_size_mm = replace_na(eop_size_mm, 0),
    eop_size = factor(eop_size, c("0", "1", "2", "3", "4", "5"), c("0-5", "5-10", "10-15", "15-20", "20-25", "25+"), ordered = TRUE), 
    sex = factor(sex, c(0, 1), c("female", "male")),
    fhp_category = factor(fhp_category,c(0, 1, 2, 3, 4, 5, 6, 7),c("0-10", "10-20", "20-30", "30-40", "40-50", "50-60", "60-70", "70-80"), ordered = TRUE),
    age_group = factor(age_group, c("2", "3", "4", "5", "6", "7", "8"), c("18-30", "31-40", "41-50", "51-60", "60+", "60+", "60+"), ordered = TRUE)) %>%
    filter(age_group != "1")
posture_data
```
Part 2:
The dataset posture_data is comprised of 9 columns. Key variables include eop_size_mm, fhp_size_mm, age, age_group, and sex. Some variables were converted to ordered factors, and others remained numeric. We have 1,219 participants included in the posture_data dataset.  
```{r}
agesex_table = posture_data %>%
filter(age_group != "1") %>%
group_by(sex, age_group) %>%
summarize (n= n()) %>%
pivot_wider(
names_from = sex, 
values_from = n
) %>%
knitr::kable ()
agesex_table

agesex_plot=
posture_data %>% 
ggplot(aes(x = age, fill = sex)) +
geom_histogram()
agesex_plot
```

Problem 2
Part 1:
```{r}
figure3 = 
ggplot(posture_data, aes(x = age, y = fhp_size_mm, na.rm = TRUE)) +
geom_point(aes(color = age_group)) + 
geom_smooth(size = .7, se = TRUE) + 
facet_grid(~sex)
figure3

figure4 = 
posture_data %>% 
mutate(enlarged_eop = as.numeric(ifelse(eop_size_mm >=10, "1", "0"))) %>% 
group_by(age_group, sex) %>% 
summarize(rate = mean(enlarged_eop)) %>% 
ggplot(aes(x = age_group, y = rate, group =sex, color = sex)) +
geom_point () +
geom_line () +
labs(
title = "Rate of enlarged EOP by Age and Sex",
x = "Age_Group",
y = "EOP Rate")
figure4

figure3+figure4
```

Part 2:
Our plots show that for both men and women the rate of EOP size is higher among age 18-30 and increased FHP is more common among age 60+.  This is in line with the findings in the original paper. 
```{r}
posture_data %>%
  ggplot(aes(x = eop_size_mm, y = fhp_size_mm, na.rm = TRUE, color = age_group)) +
         geom_point () + facet_grid (sex ~ age_group) + labs (title = "2x5 Panel" , 
    x = "eop_size_mm" , 
    y = "fhp_size_mm")

```

Problem 3
Part 1:
Yes, a few groups have 3-4 more than cited in the paper, but in general they are consistent among age groups. 
```{r}
posture_data %>%
  group_by(age_group) %>%
  summarize(n_obs = n())
```
Part 2:
Yes, our findings are consistent with the above mean and standard deviations for males and females cited in the paper. 
```{r}
mean(pull(filter(posture_data,sex == "female"), fhp_size_mm), na.rm = TRUE)
sd(pull(filter(posture_data,sex == "female"), fhp_size_mm), na.rm = TRUE)
mean(pull(filter(posture_data,sex == "male"), fhp_size_mm), na.rm = TRUE)
sd(pull(filter(posture_data,sex == "male"), fhp_size_mm), na.rm = TRUE)
```

Part 3: 
EEOP is an enlarged external occipital protuberance greater . We can use the variable eop_size_mm to determine how many of our subjects are above the 10mm specifiation for EEOP.  
The prevalence from the paper is 33% and we find that is it 32%. This is similar enough to be consistent. 
```{r}
posture_data %>%
 mutate(eeop = as.numeric(ifelse(eop_size_mm > 10, "1", "0"))) %>%
 summarize (eeop_rate = mean(eeop))
```

Part 4:
Yes, we see in the below plot that there is a greater frequency of increased FHP size in the age group 60+. 
```{r}
posture_data %>%
 filter(fhp_size_mm > 40) %>%
 group_by(age_group) %>%
 summarize(n_obs = n()) %>%
 ggplot(aes(x = age_group, y = n_obs, color = age_group)) +
geom_bar(stat = "identity") +
 geom_text(aes(label = n_obs), vjust = 2) + labs(title = "Frequency >40 FHP",
   x = "age group",
   y = "frequency >40 FHP")
```

Problem 4
The original report concludes that there is a 33% prevalance of EEOP in the population, that those aged 60+ have increased FHP size, and that those aged 18-30 have a higher rate of EEOP than other age groups due to increased mechanical load from using hand-held devices.  While the results in our analysis conincide with the results found in the paper, we do not agree with the causal effect that is demonstrated by the authors.  The methods in which the participants were recruited for the age group of 18-30 resulted in a potential bias which was erroneously interpreted by the authors. I do not think that this data is sufficient to suggest that cell phones are causing horn growth, but do think it should be further investigated. In order to make this causal assumption of hand-held devices and increaed prevalance of EEOP I would want detailed information on participant technology use, and lifestyle variables to coincide with eop and fhp sizes.  




